---
title: "training_synthetic_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{training_synthetic_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(spect)
```

# Generating unit test data

It can be useful to create a data set with a known distribution for testing novel modeling approaches. In this case, the sample data set generated is used for unit testing the spect package.

```{r}

rng_seed <- 42
set.seed(rng_seed)

syn_data <- create_synthetic_data(sample_size=2500,
                      censor_percentage = 0.1,
                      perturbation_shift = 6)
                      

source_data <- select(syn_data, -c(baseline_time_to_cancel, perturbed_baseline))

predict_data <- source_data[1:10,]
modeling_data <- source_data[11:nrow(source_data),]
```

Training the model then becomes a straightforward call to spect_train.

```{r}

event_indicator_var <- "cancel_event_detected"
survival_time_var <- "total_months"
obs_window <- 48
alg="glm"

result <- spect_train(model_algorithm=alg, modeling_data=modeling_data,
                      event_indicator_var=event_indicator_var,
                      survival_time_var=survival_time_var,
                      obs_window=obs_window, use_parallel=FALSE)
```

All of the results are collected and serialized for use in unit tests. This data is available in the testthat/testdata/helper_train_data.rds file.

```{r}

spect_test_data <- mget(c("event_indicator_var", "survival_time_var",
                          "obs_window", "alg", "syn_data", "source_data",
                          "predict_data", "modeling_data", "result", "rng_seed"))
                          
saveRDS(object=spect_test_data, file="helper_train_data.rds")
```
